```{r}
library(ggplot2)
library(tidyverse)
library(Seurat)
library(slingshot)
library(tradeSeq)
library(RColorBrewer)
library(ggpubr)
```

# load data
```{r}
Neoblast.combined <- readRDS(file = "../01.rds_files/Neoblast.combined.rds")
```

# assign cell clusters to cell types
```{r}
Neoblast.combined$seurat_clusters <- Neoblast.combined$integrated_snn_res.3

Neoblast.combined@meta.data <- Neoblast.combined@meta.data %>%
    mutate(newCell_Type = case_when(
        seurat_clusters %in% c(2, 4, 7, 10, 11, 30, 33, 34) ~ "SSCs-1", # alpha
        seurat_clusters %in% c(5, 9, 13, 15) ~ "SSCs-2", # beta
        seurat_clusters %in% c(3, 8, 16, 17, 22, 40) ~ "SSCs-5", # gamma 1
        seurat_clusters %in% c(18, 27) ~ "SSCs-6", # gamma 2
        seurat_clusters %in% c(14, 24) ~ "SSCs-7", # delta
        seurat_clusters %in% c(1, 29) ~ "SSCs-3", # 'SSCs-epsilon',
        seurat_clusters %in% c(12) ~ "SSCs-4", # kappa
        seurat_clusters %in% c(6, 20) ~ "S1",
        seurat_clusters %in% c(21) ~ "SSCs-8", # nu
        seurat_clusters %in% c(0, 26, 31) ~ "S1 progeny",
        seurat_clusters %in% c(32) ~ "Early Vo",
        seurat_clusters %in% c(42, 38) ~ "Late Vo",
        seurat_clusters %in% c(36) ~ "Mehlis gland",
        seurat_clusters %in% c(19, 28, 35) ~ "GSCs",
        seurat_clusters %in% c(23, 25) ~ "GSC progeny",
        seurat_clusters %in% c(37) ~ "Female gametophyte",
        seurat_clusters %in% c(39, 41) ~ "Male gametophyte"
    ))

Neoblast.combined@meta.data$newCell_Type <- factor(as.character(Neoblast.combined@meta.data$newCell_Type),
    levels = c(
        "SSCs-1",
        "SSCs-2",
        "SSCs-3",
        "SSCs-4",
        "SSCs-5",
        "SSCs-6",
        "SSCs-7",
        "SSCs-8",
        "GSCs",
        "GSC progeny",
        "Female gametophyte",
        "Male gametophyte",
        "S1",
        "S1 progeny",
        "Early Vo",
        "Late Vo",
        "Mehlis gland"
    )
)
```


# FigS7a
```{r fig.height=5, fig.width=12}
col_pal <- col_vectors[1:length(unique(Neoblast.combined@meta.data$newCell_Type))]
FigS7a <- create_numbered_umap(seurat_obj = Neoblast.combined, split_by = "Sex", col_vectors = col_pal, pointSize = 0.01, labelSize = 10)
ggsave("../02.figures/FigS7a.pdf", width = 12, height = 5, plot = FigS7a)
```

# FigS7b/7c
```{r fig.height=4.5, fig.width=10}
DefaultAssay(Neoblast.combined) <- "RNA"
Idents(Neoblast.combined) <- Neoblast.combined$newCell_Type


ids <- c(
    "Sjc-0000924", "Sjc-0005962", "Sjc-0008047",
    #' Sjc-0001581',
    "Sjc-0001581", "Sjc-0007059", "Sjc-0008087", "Sjc-0001087", # δ'
    "Sjc-0009220", "Sjc-0001001", "Sjc-0006063",
    "Sjc-0001890",
    "Sjc-0005317",
    "Sjc-0005761", "Sjc-0006155", # epsilon
    "Sjc-0001481", "Sjc-0005786",
    "Sjc-0000674", "Sjc-0004091",
    "Sjc-0006622", "Sjc-0009231", "Sjc-0008219",
    "Sjc-0004774", "Sjc-0007561", # GSC
    "Sjc-0004512", "Sjc-0003040",
    "Sjc-0001604", "Sjc-0000843",
    "Sjc-0000108", "Sjc-0002828",
    "Sjc-0003451", "Sjc-0007461",
    "Sjc-0006798", "Sjc-0006800", "Sjc-0000278", "Sjc-0003576", "Sjc-0000795"
)



# ID2NAME_new <- ID2NAME
ID2NAME_new$gene_name <- ID2NAME_new$new_gene_name
original_gene_ids <- ids
custom_gene_names <- convert_id_name(ids, "id", ID2NAME_new)
names(custom_gene_names) <- original_gene_ids

# 首先检查数据的分布
plot_data <- Seurat::DotPlot(Neoblast.combined, features = original_gene_ids, scale = FALSE)$data
quantile(plot_data$avg.exp, c(0.05, 0.25, 0.5, 0.75, 0.95))



Fig_dotplot_cellmarker.f <- Seurat::DotPlot(
    subset(Neoblast.combined, Sex == "Female"),
    features = original_gene_ids,
    scale = TRUE
) +
    theme(
        axis.text = element_text(size = 10, hjust = 1),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.border = element_rect(fill = NULL, linewidth = 1, colour = "black")
    ) +
    scale_colour_gradientn(
        colours = c("grey", rev(brewer.pal(n = 11, name = "Spectral"))[-1]),
        limits = c(-2, 2), # 限制颜色范围，增强对比度
        oob = scales::squish, # 超出范围的值被压缩到边界
        name = "Log Expression"
    ) +
    scale_x_discrete(labels = custom_gene_names) +
    xlab("") + ylab("")


Fig_dotplot_cellmarker.m <- Seurat::DotPlot(
    subset(Neoblast.combined, Sex == "Male"),
    features = original_gene_ids,
    scale = TRUE
) +
    theme(
        axis.text = element_text(size = 10, hjust = 1),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.border = element_rect(fill = NULL, linewidth = 1, colour = "black")
    ) +
    scale_colour_gradientn(
        colours = c("grey", rev(brewer.pal(n = 11, name = "Spectral"))[-1]),
        limits = c(-2, 2), # 限制颜色范围，增强对比度
        oob = scales::squish, # 超出范围的值被压缩到边界
        name = "Log Expression"
    ) +
    scale_x_discrete(labels = custom_gene_names) +
    xlab("") + ylab("")

p <- Fig_dotplot_cellmarker.f + Fig_dotplot_cellmarker.m

ggsave("../02.figures/FigS7bc.pdf", width = 20, height = 6, plot = p)
```



# FigS11a
```{r fig.height=6, fig.width=8}
library(reshape2)
library(ggpubr)
library(ggalluvial)

Idents(Neoblast.combined) <- Neoblast.combined@meta.data$Cell_Type
df <- Neoblast.combined@meta.data %>%
    dplyr::select("sample", "Cell_Type") %>%
    table()
df <- as.data.frame(prop.table(df, 1) * 100)
df$sample <- factor(df$sample, levels = c(
    "26dpi_F", "22dpi_F", "18dpi_F",
    "18dpi_M", "22dpi_M", "26dpi_M"
))

df$Sex <- if_else(grepl("F", df$sample), "Female", "Male")

FigS11a <- map(split(df, df$Sex), function(d) {
    ggplot(
        d,
        aes(
            x = sample, stratum = Cell_Type, alluvium = Cell_Type,
            y = Freq,
            fill = Cell_Type, label = Cell_Type
        )
    ) +
        scale_x_discrete(expand = c(.1, .1)) +
        scale_fill_manual(values = col_vectors[1:length(unique(Neoblast.combined@meta.data$Cell_Type))]) +
        geom_flow() +
        geom_stratum(
            stat = "alluvium",
            color = "darkgray"
        ) +
        theme(legend.position = "right", axis.line = element_blank(), panel.background = element_blank()) +
        ggtitle("Frequency of cells in different conditions")
}) %>% patchwork::wrap_plots(ncol = 2, guides = "collect")

ggsave("../02.figures/FigS11a.pdf", width = 8, height = 6, plot = FigS11a)
FigS11a
```


# FigS11b
```{r fig.height=4, fig.width=6}
FigS11b <- barplot_per(Neoblast.combined,
    colValues = c(
        "#B2182B", "#EF8A62", "#FDDBC7", "#D1E5F0", "#67A9CF", "#2166AC"
    )
    # ,levels = c('26dpi_F','22dpi_F','18dpi_F','18dpi_M', '22dpi_M','26dpi_M')
) + xlab("") + theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.2, face = "bold"))

ggsave("../02.figures/FigS11b.pdf", width = 6, height = 4, plot = FigS11b)
FigS11b
```



# Sub-clustering analysis of vitelline lineage
```{r}
library(gam)
library(slingshot)
library(SingleCellExperiment)
library(scater)
library(ggpubr)
DefaultAssay(Neoblast.combined.f) <- "RNA"

saveRDS(Neoblast.combined.f.degs, "../01.rds_files/Neoblast.combined.f.degs.rds")

seu_vo_lineage <- Neoblast.combined.f[, Neoblast.combined.f$newCell_Type %in%
    c("S1", "S1 progeny", "Early Vo", "Late Vo")]
```

# Fine-cell annotation of vitelline lineage
```{r}
seu_vo_lineage@meta.data <- seu_vo_lineage@meta.data %>%
    mutate(fineCell_Type = case_when(
        seurat_clusters %in% c(20) ~ "S1_0",
        seurat_clusters %in% c(6) ~ "S1_1",
        seurat_clusters %in% c(31) ~ "S1p_0",
        seurat_clusters %in% c(26) ~ "S1p_1",
        seurat_clusters %in% c(0) ~ "S1p_2",
        seurat_clusters %in% c(32) ~ "EVo",
        seurat_clusters %in% c(42) ~ "LVo_0",
        seurat_clusters %in% c(38) ~ "LVo_1"
    ))
seu_vo_lineage@meta.data$fineCell_Type <- factor(as.character(seu_vo_lineage@meta.data$fineCell_Type),
    levels = c(
        "S1_0", "S1_1", "S1p_0",
        "S1p_1", "S1p_2", "EVo",
        "LVo_0", "LVo_1"
    )
)
```


# FigS9a

```{r fig.height=4, fig.width=5}
library(ggpubr)
library(RColorBrewer)
library(tidyverse)

# plot_slingshot(SlingshotDataSet(sce),show_arrow = T,curve_range = c(0.15, 0.95))

pumap <- create_numbered_umap(seu_vo_lineage, group_by = "fineCell_Type", col_vectors = col_vector)

ggsave("../02.figures/FigS9a.pdf", width = 5, height = 4, plot = pumap)
```




# Dotplot of select marker genes
```{r}
filter_markers <- function(seumarker, ID2NAME) {
    select_Markers <- seumarker %>%
        mutate(GeneName = ID2NAME[gene, "Gene_Name"]) %>%
        filter(!(grepl("^ND|^C|^ATP", gene))) %>%
        filter(!(grepl("^Sjc", GeneName))) %>%
        filter(!(grepl("^R[P][SL]", toupper(GeneName)))) %>%
        group_by(cluster) %>%
        top_n(n = -20, wt = p_val_adj) %>%
        group_by(cluster) %>%
        dplyr::top_n(n = 10, wt = avg_log2FC) %>%
        top_n(n = 5, wt = avg_log2FC) %>%
        distinct(gene, .keep_all = T)
    return(select_Markers)
}


neo_cellmarker_f %>%
    mutate(Note = ID2NAME[gene, "Note"]) %>%
    filter(!(grepl("^ND|^C|^ATP", gene))) %>%
    dplyr::filter(cluster %in% c(42))

select_Markers <- filter_markers(neo_cellmarker_f, ID2NAME)

select_Markers


vo_cellmarker_ids <- neo_marker_cosg$names[, colnames(neo_marker_cosg$names) %in% unique(seu_vo_lineage$seurat_clusters)] %>%
    stack() %>%
    dplyr::filter(!grepl("^ND|^C|^ATP", values)) %>%
    dplyr::mutate(cls = factor(ind, levels = c(20, 6, 31, 26, 0, 32, 42, 38))) %>%
    dplyr::arrange(cls) %>%
    dplyr::group_by(cls) %>%
    dplyr::slice_head(n = 5) %>%
    pull(values)


neo_cellmarker_f_subvo <- neo_cellmarker_f[neo_cellmarker_f$cluster %in% unique(seu_vo_lineage$seurat_clusters), ]


subvo_cls <- c(
    "S1_0", "S1_1", "S1p_0",
    "S1p_1", "S1p_2",
    "EVo", "LVo_0", "LVo_1"
)
names(subvo_cls) <- c(20, 6, 31, 26, 0, 32, 42, 38)
neo_cellmarker_f_subvo$cluster <- subvo_cls[as.character(neo_cellmarker_f_subvo$cluster)]

neo_cellmarker_f_subvo$cluster <- factor(neo_cellmarker_f_subvo$cluster, levels = c(
    "S1_0", "S1_1", "S1p_0",
    "S1p_1", "S1p_2", "EVo",
    "LVo_0", "LVo_1"
))

neo_cellmarker_f_subvo <- neo_cellmarker_f_subvo %>%
    dplyr::filter(p_val_adj < 0.05) %>%
    dplyr::arrange(cluster)


writexl::write_xlsx(
    split(neo_cellmarker_f_subvo, neo_cellmarker_f_subvo$cluster),
    "../02.figures/tables/vo_cellmarker.xlsx"
)
```


# FigS9b
```{r fig.height=5, fig.width=9}
ids <- c(
    "Sjc-0000108", "Sjc-0000278", "Sjc-0002805",
    "Sjc-0001662", "Sjc-0001005", "Sjc-0005295", "Sjc-0005365", "Sjc-0006610",
    "Sjc-0004280", "Sjc-0002828", "Sjc-0002171", "Sjc-0006066", "Sjc-0006745",
    "Sjc-0003538", "Sjc-0003891", "Sjc-0000177", "Sjc-0006030", "Sjc-0001878",
    "Sjc-0009101", "Sjc-0004269", "Sjc-0000014", "Sjc-0001339", "Sjc-0000087",
    "Sjc-0004097", "Sjc-0009196", "Sjc-0005816", "Sjc-0005796", "Sjc-0003393",
    "Sjc-0005021", "Sjc-0001886", "Sjc-0001888", "Sjc-0005034",
    "Sjc-0007461", "Sjc-0006798", "Sjc-0006799",
    "Sjc-0006673",
    "Sjc-0004903", "Sjc-0000117", "Sjc-0004267", "Sjc-0000244"
)

ID2NAME_new <- ID2NAME
ID2NAME_new$gene_name <- ID2NAME$new_gene_name

original_gene_ids <- ids
custom_gene_names <- convert_id_name(ids, "id", ID2NAME_new)
names(custom_gene_names) <- original_gene_ids



Idents(seu_vo_lineage) <- seu_vo_lineage$fineCell_Type
DefaultAssay(seu_vo_lineage) <- "RNA"
dp <- DotPlot(seu_vo_lineage, features = ids) + RotatedAxis() +
    theme_pubr(border = T) + RotatedAxis() +
    theme(
        axis.text = element_text(size = 10, hjust = 1),
        panel.border = element_rect(fill = NULL, linewidth = 1, colour = "black")
    ) +
    scale_colour_gradientn(colours = c("grey", rev(brewer.pal(n = 11, name = "Spectral"))[-1])) +
    RotatedAxis() +
    scale_x_discrete(labels = custom_gene_names) +
    xlab("") + ylab("")
dir.create("../02.figures/vo_lineage/")
ggsave(filename = "../02.figures/vo_lineage/FigS9b.pdf", width = 9, height = 5, plot = dp)
```


# Trajectory_inference using Slingshot

### 1) Choose start cell states
```{r}
DefaultAssay(seu_vo_lineage) <- "RNA"
Idents(seu_vo_lineage) <- seu_vo_lineage$fineCell_Type
VlnPlot(seu_vo_lineage, features = c("Sjc-0000108", "Sjc-0009220", "Sjc-0000924"))
```
CytoTRACE
```{r}
# FigS9e
expr <- seu_vo_lineage@assays$RNA@counts[rowSums(seu_vo_lineage@assays$RNA@counts) > 0, ]
res <- CytoTRACE::CytoTRACE(as.matrix(expr))
cls_l <- as.character(seu_vo_lineage@meta.data$fineCell_Type)
names(cls_l) <- rownames(seu_vo_lineage@meta.data)
CytoTRACE::plotCytoTRACE(
    cyto_obj = res,
    # gene = 'Sjc-0005962',
    phenotype = cls_l, # 细胞类型注释
    emb = Seurat::Embeddings(seu_vo_lineage, reduction = "umap"), # 是否提供细胞降维坐标
    outputDir = "../02.figures/" # 图片储存路径
)
```
### 2) Run Slingshot
```{r}
DefaultAssay(seu_vo_lineage) <- "integrated"

sce <- as.SingleCellExperiment(seu_vo_lineage, assay = "integrated")
sce <- slingshot(
    sce,
    reducedDim = "UMAP",
    clusterLabels = "fineCell_Type",
    start.clus = "S1_0", end.clus = c("LVo_0", "LVo_1"),
    stretch = 1,
    df = 5,
    shrink = T
)
```

```{r}
saveRDS(sce, "../01.rds_files/slingshot_sce_voLinegae.rds")
sce <- readRDS("../01.rds_files/slingshot_sce_voLinegae.rds")
```


### 3) Visualize
# FigS9cd
```{r fig.height=6, fig.width=8}
p <- plot_trajectory_on_background(sce, seu_vo_lineage, lineage_idx = 1, line_width = 0.5, point_size = 0.05, curve_range = c(0.15, 1)) /
    plot_trajectory_on_background(sce, seu_vo_lineage, lineage_idx = 2, line_width = 0.5, point_size = 0.05, curve_range = c(0.13, 0.85))

p <- p & theme_pubr(border = T, legend = "right")
ggsave("../02.figures/FigS9cd.pdf", width = 8, height = 6, plot = p)
```


### 4) Tradeseq
```{r}
library(tradeSeq)
# seurat :seu_vo_lineage
# slingshot : sce
counts <- GetAssayData(object = seu_vo_lineage, assay = "RNA", slot = "counts")

set.seed(5)
icMat <- evaluateK(
    counts = counts, sds = crv, k = 3:10,
    nGenes = 200, verbose = T, plot = TRUE
)
```


```{r}
job::job(
    {
        library(slingshot)
        library(tradeSeq)
        set.seed(6)
        crv <- SlingshotDataSet(sce)
        sce.new <- fitGAM(counts = counts, sds = crv, nknots = 6, verbose = FALSE)
        saveRDS(sce.new, "~/01.rds_files/tradeseq_voLinegae_f.rds")
    },
    import = list(counts, sce)
)

sce.new <- readRDS("~/01.rds_files/tradeseq_voLinegae_f.rds")
```

```{r}
table(rowData(sce.new)$tradeSeq$converged)
```

# associationTest
```{r}
assoRes <- associationTest(sce.new)
head(assoRes)
assoRes %>% dplyr::filter(pvalue < 0.05)
```


```{r}
assocRes_lineages <- associationTest(sce.new, lineages = TRUE, l2fc = log2(2))
```


```{r}
plot_trend_glist <- function(glist, sce = sce.new, counts = counts) {
    p_trend <- lapply(glist$gene, function(x) {
        p <- plotSmoothers(sce, counts, gene = x) +
            ggtitle(label = ID2NAME[x, ]$new_gene_name, subtitle = ID2NAME[x, ]$Note)
        p
    })
    names(p_trend) <- glist$gene
    return(p_trend)
}
```


```{r}
library(ComplexHeatmap)
library(circlize)
library(ClusterGVis)
library(org.Sjaponicum.eg.db)
```


```{r}
mat <- getGroupMat(sce, counts, select_genes = startRes_glist$gene)
getClusters(obj = as.data.frame(na.omit(smoothMatrix(mat))))
mat <- as.data.frame(na.omit(smoothMatrix(mat)))
rownames(mat) <- revID(rownames(mat))
ck <- clusterData(
    obj = mat,
    scaleData = T,
    cluster.method = "kmeans",
    cluster.num = 4
)
```


```{r}
ptime <- seq(0, 1, length.out = 100)
topha <- HeatmapAnnotation(
    Pseudotime = ptime,
    col = list(Pseudotime = colorRamp2(seq(0, 1, 1 / (length(cList$`Blue-Yellow-Red`) - 1)), cList$`Blue-Yellow-Red`))
)
```


```{r}
GO_objList <- lapply(ck$cluster.list,
    SJDB::getGO,
    universeIDs = revID(rownames(seu_vo_lineage)), version = "V3"
)

ko_c1 <- SJDB::getKEGG(
    geneid = ck$cluster.list$C2,
    universeIDs = revID(rownames(seu_vo_lineage)), version = "V3"
)

# getGO(ck$cluster.list$C6, universeIDs = revID(rownames(seu_vo_lineage)), version = "V3", qCutoff = 1)

GO_list <- lapply(ck$cluster.list, function(x) {
    go <- clusterProfiler::enrichGO(
        gene = x, OrgDb = "org.Sjaponicum.eg.db",
        keyType = "GENEID", ont = "BP", pvalueCutoff = 1,
        pAdjustMethod = "BH", qvalueCutoff = 1,
        minGSSize = 10, maxGSSize = 500, readable = FALSE
    )
    go
})

C5 <- clusterProfiler::enrichGO(
    gene = ck$cluster.list$C5, OrgDb = "org.Sjaponicum.eg.db",
    keyType = "GENEID", ont = "BP", pvalueCutoff = 1,
    pAdjustMethod = "BH", qvalueCutoff = 1,
    minGSSize = 10, maxGSSize = 500, readable = FALSE
)
```



```{r}
revigo <- function(GO_obj,
                   n = 50,
                   p_cutoff = 0.01,
                   version = "V3") {
    library(rrvgo)
    library(clusterProfiler)
    if (version == "V3") {
        OrgDB <- "org.Sjaponicum.eg.db"
    } else if (version == "V4") {
        OrgDB <- "org.SjaponicumV4.eg.db"
    } else {
        print("Please provide genome version V3 or V4!")
    }

    bp <- as.data.frame(GO_obj)

    if (dim(bp)[1] >= 1) {
        bp_top <- bp %>%
            dplyr::filter(pvalue < p_cutoff)

        tryCatch(
            {
                simMatrix <- calculateSimMatrix(bp_top$ID,
                    keytype = "GENEID",
                    orgdb = OrgDB,
                    ont = "BP",
                    method = "Rel"
                )

                scores <- setNames(-log10(bp_top$qvalue), bp_top$ID)
                reducedTerms <- reduceSimMatrix(simMatrix,
                    scores,
                    threshold = 0.7, keytype = "GENEID",
                    orgdb = OrgDB
                )

                bp_top <- bp_top[bp_top$ID %in% reducedTerms$parent, ]
            },
            error = function(e) {
                message("Caught an error: ", e$message)
                bp_top <- bp
            }
        )
    } else {
        bp_top <- NULL
        message("No enrichment GO termss")
    }
    return(bp_top)
}


revigo_list <- lapply(GO_list, revigo)


revigo_list
```


#### *associationTest
```{r}
assoRes <- associationTest(sce.new, lineages = TRUE)
assoRes_glist <- assoRes %>%
    dplyr::filter(pvalue_1 < 1e-5) %>%
    dplyr::mutate(
        gene = rownames(.),
        gene_name = ID2NAME[gene, "gene_name"],
        Note = ID2NAME[gene, "Note"]
    )

assoRes_glist
```



```{r}
mat <- getGroupMat(sce, counts, select_genes = assoRes_glist$gene)
getClusters(obj = as.data.frame(na.omit(smoothMatrix(mat))))
```


```{r}
mat <- as.data.frame(na.omit(smoothMatrix(mat)))
rownames(mat) <- revID(rownames(mat))
ck <- clusterData(
    obj = mat,
    scaleData = F,
    cluster.method = "kmeans",
    cluster.num = 6
)
```


```{r}
bin_celltypes <- getBinCellTypes_fixed(sce, Pseudotime = "slingPseudotime_1", celltype_col = "fineCell_Type")

cell_type_colors <- col_vector[1:length(levels(colData(sce)$fineCell_Type))]
names(cell_type_colors) <- levels(colData(sce)$fineCell_Type)


ptime <- seq(0, 1, length.out = 100)

topha <- HeatmapAnnotation(
    Pseudotime = ptime,
    CellType = bin_celltypes,
    col = list(
        Pseudotime = colorRamp2(seq(0, 1, 1 / (length(cList$`Blue-Yellow-Red`) - 1)), cList$`Blue-Yellow-Red`),
        CellType = cell_type_colors[c(1, 2, 3, 4, 6, 7)]
    )
)
```


# FigS9g
```{r fig.height=6.5, fig.width=10.5}
GO_list.vo <- lapply(ck$cluster.list, function(x) {
    go <- clusterProfiler::enrichGO(
        gene = x, OrgDb = "org.Sjaponicum.eg.db",
        keyType = "GENEID", ont = "BP", pvalueCutoff = 1,
        pAdjustMethod = "BH", qvalueCutoff = 1,
        minGSSize = 10, maxGSSize = 500, readable = FALSE
    )
    return(revigo(go))
})

termanno.vo <- df.vo %>%
    dplyr::group_by(cluster) %>%
    dplyr::slice_min(order_by = pvalue, n = 4) %>%
    dplyr::ungroup() %>%
    dplyr::select(cluster, Description, pvalue) %>%
    dplyr::rename(id = cluster, term = Description)
termanno.vo$pvalue[termanno.vo$pvalue < 1e-10] <- 1e-10


saveRDS(ck, "../02.figures/volineage_ck.rds")

pdf("../02.figures/FigS9g.pdf", width = 10.2, height = 5.5)
visCluster(
    object = ck,
    plot.type = "both",
    HeatmapAnnotation = topha,
    annoTerm.data = termanno.vo,
    term.text.limit = c(9, 12),
    cluster.order = c(1, 3, 6, 5, 4, 2),
    add.sampleanno = T
)
dev.off()
```



# Export
```{r}
saveRDS(seu_vo_lineage, "../01.rds_files/seu_vo_lineage.rds")
saveRDS(sce, "../01.rds_files/sce_vo_lineage.rds")
```


```{r}
plot_trajectory_on_background <- function(sce,
                                          full_seurat_obj,
                                          lineage_idx = 1,
                                          point_size = 0.5,
                                          arrow_size = 0.1,
                                          line_width = 1.0) {
    library(ggplot2)
    library(RColorBrewer)
    library(dplyr)
    library(SingleCellExperiment)

    # 1. 准备绘图数据框
    # 从完整的Seurat对象中获取所有细胞的UMAP坐标
    all_coords <- as.data.frame(Embeddings(full_seurat_obj, reduction = "umap"))
    all_coords$cell_id <- rownames(all_coords)

    # 从sce对象中获取轨迹细胞的伪时间
    traj_cells <- colnames(sce)
    pdt <- slingPseudotime(sce)

    # 检查并选择正确的lineage
    if (ncol(pdt) > 1 && lineage_idx > ncol(pdt)) {
        warning("lineage_idx is out of bounds. Using the first lineage.")
        lineage_idx <- 1
    }

    pdt_df <- data.frame(
        cell_id = traj_cells,
        pseudotime = pdt[, lineage_idx]
    )

    # 2. 合并数据框
    # 将所有细胞的坐标与轨迹细胞的伪时间合并
    # 非轨迹细胞的 pseudotime 值将为 NA
    plot_df <- left_join(all_coords, pdt_df, by = "cell_id")

    # 3. 创建ggplot对象
    # 首先绘制所有点，NA值（背景细胞）将被设置为灰色
    p <- ggplot(plot_df, aes(x = UMAP_1, y = UMAP_2)) +
        geom_point(aes(color = pseudotime), size = point_size) +
        # 使用 scale_color_gradientn 来更好地控制颜色，并将NA值设为灰色
        scale_color_gradientn(
            colors = rev(brewer.pal(11, "Spectral")[-6]),
            na.value = "grey80", # <-- 这里是关键！
            name = "Pseudotime"
        ) +
        theme_classic() +
        labs(title = paste("Lineage", lineage_idx, "on Full UMAP")) +
        coord_equal() # 保持UMAP的纵横比

    # 4. 添加轨迹曲线
    curves < -slingCurves(sce, as.df = TRUE)
    if ("Lineage" %in% colnames(curves)) {
        curves_selected <- curves[curves$Lineage == lineage_idx, ]
    } else {
        curves_selected <- curves
    }

    if (nrow(curves_selected) > 0) {
        arrow_style <- arrow(type = "closed", length = unit(arrow_size, "inches"))

        p <- p + geom_path(
            data = curves_selected %>% arrange(Order),
            aes(x = UMAP_1, y = UMAP_2), # group aesthetic is not needed here
            color = "black", size = line_width,
            arrow = arrow_style
        )
    }

    return(p)
}
```


```{r}
plot_slingshot <- function(sds, show_arrow) {
    test <- data.frame(reducedDims(sds))
    if (length(sds@lineages) > 1) {
        pdt_agg1 <- apply(slingPseudotime(sds)[, c(1:length(sds@lineages))], 1, function(t) mean(t, na.rm = T))
    } else {
        pdt_agg1 <- slingPseudotime(sds)
    }
    test$pseudotime <- pdt_agg1
    p <- ggscatter(test, x = "UMAP_1", y = "UMAP_2", color = "pseudotime", size = 0.8) +
        gradient_color(palette = rev(brewer.pal(11, "Spectral")[-6])) +
        theme_pubr(border = T)

    curves <- slingCurves(sds, as.df = TRUE)
    arrow_style <- arrow(type = "closed", length = unit(0.1, "inches"))
    if (show_arrow) {
        p <- p + geom_path(
            data = curves %>% arrange(Order),
            aes(group = Lineage), show.legend = T, arrow = arrow_style
        )
    }
    p
}
```


```{r fig.height=5, fig.width=4.5}
plot_slingshot(SlingshotDataSet(sce), show_arrow = T)
```



```{r}
Idents(seu_vo_lineage) <- seu_vo_lineage@meta.data$newCell_Type
marker_cosg <- COSG::cosg(
    seu_vo_lineage,
    groups = "all",
    assay = "RNA",
    slot = "data",
    mu = 1,
    n_genes_user = 100
)
DotPlot(seu_vo_lineage, features = stack(head(marker_cosg$names, 5))$value) & RotatedAxis()
```






```{r}
library(viridis)
ptplot <- function(sce, lineage = "slingPseudotime_1", m, n) {
    p <- as.data.frame(colData(sce)[, lineage])
    colnames(p) <- "slingpt"
    p$new <- p$slingpt
    p$new[which(p$slingpt < m)] <- NA
    p$new[which(p$slingpt > n)] <- NA

    # 使用viridis调色板
    color_palette <- viridis(100)
    plotcol <- color_palette[cut(p$new, breaks = 100)]
    plotcol <- ifelse(is.na(plotcol), "grey", plotcol)

    plot(reducedDims(sce)$UMAP, col = plotcol, pch = 16, asp = 1)
}
```





```{r}
seu_subset <- subset(Neoblast.combined.f, newCell_Type %in% c(
    "S1",
    "S1 progeny",
    "Early Vo",
    "Late Vo"
))


seu_subset <- FindVariableFeatures(seu_subset)
seu_subset <- ScaleData(seu_subset)
seu_subset <- RunPCA(seu_subset)
seu_subset <- RunUMAP(seu_subset, dims = 1:15) # 使用适合子集的PC数量


# 3. 在子集上运行Slingshot (Run Slingshot on the subset)
umap_subset <- Embeddings(seu_subset, "umap")
clusters_subset <- Idents(seu_subset)

sds_subset <- slingshot(umap_subset,
    clusterLabels = clusters_subset,
    start.clus = "S1"
)

DimPlot(seu_subset)

plot_slingshot(SlingshotDataSet(sds_subset))
```



```{r  fig.height=8, fig.width=12}
DimPlot(seu_neo, group.by = c("newCell_Type", "seurat_clusters", "phase"), split.by = "sample") & scale_color_manual(values = col_pal)
```
## COSG
```{r}
Idents(seu_neo) <- seu_neo@meta.data$seurat_clusters
marker_cosg <- COSG::cosg(
    seu_neo,
    groups = "all",
    assay = "RNA",
    slot = "data",
    mu = 1,
    n_genes_user = 100
)
```



```{r fig.height=15, fig.width=25}
ids <- c(
    "Sjc-0000108", # S1
    "Sjc-0003451", # S1 progeny/early
    "Sjc-0007461", # late vitellocytes
    "Sjc-0001887", # mature vitellocytes
    "Sjc-0006753",
    "Sjc-0006798",
    "Sjc-0006800",
    "Sjc-0006622",
    "Sjc-0000278",
    "Sjc-0000795",
    "Sjc-0003576"
)

FeaturePlot(seu_neo, features = ids, min.cutoff = "q1", max.cutoff = "q99") & NoAxes() & NoLegend()
```


```{r fig.height=25, fig.width=25}
cell.markers <- stack(head(marker_cosg$names, 2)) %>% pull(values)

FeaturePlot(seu_neo, features = cell.markers, min.cutoff = "q1", max.cutoff = "q99") & NoAxes() & NoLegend()
```

### Slingshot
```{r fig.height=5, fig.width=5}
library(ggpubr)
library(RColorBrewer)
library(dplyr)
Idents(seu_neo) <- seu_neo$newCell_Type
sce <- as.SingleCellExperiment(seu_neo, assay = "RNA")

sce_neo <- slingshot(
    sce,
    reducedDim = "UMAP",
    clusterLabels = "newCell_Type",
    start.clus = "S1",
    stretch = 1,
    # allow_breaks,
    thresh = 0.05
)

sds_neo <- SlingshotDataSet(sce_neo)

plot_slingshot(sds_neo, show_arrow = T)
```


```{r}
FeaturePlot(subset(seu_neo, `Sjc-0000108` <= 0.0001), features = c("Sjc-0000278"), min.cutoff = "q1", max.cutoff = "q99", ncol = 3) & NoAxes() & NoLegend()
```


# RNA velocity
```{r}
source("R/Velocity.R")
save_filesFORscVelo(Neoblast.combined, filename = "Neoblast.combined", baseDIR = "../04.velocity/")
```
